Large-eddy simulations of the ice-shelf oceanic boundary layer

### 2021-04-19 main.tex

\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{A_suppl_fig}

@@

There are two sets of simulations which have a base case in common. The base case has low far-field thermal driving of 0.1 $^{\circ}$C and a relatively high slope for Antarctic ice shelves of 1$^{\circ}$ and vigorous far-field geostrophic flow of 20 cm s$^{-1}$ to the [XX east]. These conditions were chosen to favor a high TKE regime, for reasons discussed in Section \ref{sec:disc}. To examine the relationship between thermal driving and melt rate, as well as turbulent flow characteristics, we vary the far-field thermal driving from 0.1 $^{\circ}$C to 0.5 $^{\circ}$C in the first set of simulations. In the second set of simulations we reduce the slope from 1$^{\circ}$ to 0.01$^{\circ}$.

@@

The simulation duration is 50 h, corresponding to a little over 4 inertial periods. By this point, the turbulent intensity as measured by the maximum vertical velocity fluctuation, is reasonably steady-state over one inertial period for the 1$^{\circ}$-slope simulations (Figure S\ref{fig:max_w2_timeseries}).

@@

% Sensitivity to thermal driving

The fluctuations in TKE with flow orientation under inertial oscillations are replicated when the mean pressure gradient is rotated, with increased TKE when flow is oriented downslope (not shown or supplemental).

>>>>

\usepackage{xr}
\include{A_suppl_fig.tex}

@@

There are two sets of simulations which have a base case in common. The base case has low far-field thermal driving of 0.1 $^{\circ}$C and a relatively high slope for Antarctic ice shelves of 1$^{\circ}$ and vigorous far-field geostrophic flow of 20 cm s$^{-1}$ generated by a 0.03 pressure gradient. These conditions were chosen to favor a high TKE regime, for reasons discussed in Section \ref{sec:disc}. To examine the relationship between thermal driving and melt rate, as well as turbulent flow characteristics, we vary the far-field thermal driving from 0.1 $^{\circ}$C to 0.5 $^{\circ}$C in the first set of simulations. In the second set of simulations we reduce the slope from 1$^{\circ}$ to 0.01$^{\circ}$.

@@

The simulation duration is 50 h, corresponding to a little over 4 inertial periods. By this point, the turbulent intensity as measured by the maximum vertical velocity fluctuation, is reasonably steady-state over one inertial period for the 1$^{\circ}$-slope simulations (Figure S\ref{fig:max_w2_timeseries}). The turbulent intensity for cases with slopes less than 1$^{\circ}$ continues to decline, as we discuss later.

@@

% Melt sensitivity to thermal driving
Melt rates decline over the course of the simulations (Figure \ref{fig:melt_timeseries}), preventing the identification of steady-state melt rates. 
This decline is in response to a co-committent increase in stratification during the course of the simulation, which decreases vertical heat fluxes by reducing vertical turbulent fluctuations (Fig. S\ref{fig:max_w2_timeseries}). We do not continue our simulations in the hope of reaching steady-state conditions because the increase in stratification reduces the turbulent length scales and necessitates higher model resolution than we have used here. The high thermal driving case approaches steady-state more rapidly than low thermal driving cases due to increased shear production of TKE from the buoyant acceleration of the IOBL.

% Melt rate relationship with inertial period
The dominant temporal frequency in melt rate is the inertial frequency; however, the melt response to those oscillations is highly nonlinear. Maximum melt rates occur when the mean flow is oriented between the up-slope direction and the Coriolis-favored direction, and minimum melt rates roughly 180$^{\circ}$ opposed to that. Melt rate fluctuations increase as thermal driving increases, which we attribute to the increasing importance of buoyancy forcing in driving a near-boundary plume and thus determining the depth-distribution of shear. During high-melt periods, the far-field flow is oriented up-slope and shear production of TKE is concentrated near the boundary. During low-melt periods, the far-field flow is oriented down-slope and shear production of TKE is concentrated a few meters away from the boundary. The vertical heat flux profile through the water column follows the profile of shear production of TKE (Fig. \ref{fig:dT_profiles}, Fig. S\ref{fig:TKE_balance}). The low-melt period is somewhat brief, roughly 1/4 of the inertial period, as the onset is delayed by the time it takes to increase near-boundary stratification. In simulations where the background pressure gradient orientation was varied, we also observe reduced TKE when the mean flow is down-slope, suggesting this relationship is not unique to inertial oscillations but is more closely related to the vertical velocity structure and the depth of maximum shear. (Fig. S\ref{fig:e_orientation}).

The time-averaged melt rates are roughly linear with thermal driving (Figure \ref{fig:melt_sensitivity}a). During high-melt periods, the thermal exchange coefficient evaluated using a parameterization, is about 0.006 for all thermal driving values, less than the constant value of 0.011 reported in Jenkins et al. (2010) and commonly-used elsewhere, due to accounting for the degree of stratification. The time-averaged value decreases somewhat in the high thermal driving cases due to pronounced periods of low melt (Figure Gamma T vs. thermal driving). 
%The constant flux layer is ~2m.

Boundary layer depths as characterized by a peak in the ratio of vertical to horizontal velocity variance range from 15 m to 25 m for the moderate to low thermal driving cases.

In these simulations we find that the destruction of TKE by the stabilizing buoyancy flux is two orders of magnitude smaller than shear production of TKE. Given that the background flow is relatively strong for ice shelf settings, this result may be only narrowly applicable. However, it is consistent with \citep{davis_nicholls} who found that shear production was an order of magnitude greater than buoyancy destruction of TKE in the IOBL below Larsen C Ice Shelf.

% Description of simulation features

\begin{figure}
    \centering
    \begin{minipage}{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/melt_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/us_cmp_dT_t.png}
    \end{minipage}%
    \caption{Thermal driving simulations. (a) evolution of melt rate (b) evolution of friction velocity}
    \label{fig:dT_timeseries}
\end{figure}

\begin{figure}
    \centering
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 4cm 0 0},clip, width=\textwidth]{Figures/ptfar_cmp_dT_40hr_tav1_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 4cm 0 0},clip, width=\textwidth]{Figures/safar_cmp_dT_40hr_tav1_z_profile.png}
    \end{minipage}
    \newline
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 7.5cm 0 0},clip, width=\textwidth]{Figures/velocity_cmp_dT_40hr_tav1_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 7.5cm 0 0},clip, width=\textwidth]{Figures/Fshear_cmp_dT_40hr_tav1_z_profile.png}
    \end{minipage}
    \newline
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 7.5cm 0 0},clip, width=\textwidth]{Figures/velocity_cmp_dT_48hr_tav1_z_profile.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 7.5cm 0 0},clip, width=\textwidth]{Figures/Fshear_cmp_dT_48hr_tav1_z_profile.png}
    \end{minipage}
    \caption{Thermal driving simulations. (a) temperature relative to far-field temperature (b) salinity relative to far-field salinity (c) velocity (d) heat flux}
    \label{fig:dT_profiles}
\end{figure}

\begin{figure}
    \centering
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 3.5cm 0 0},clip,width=\textwidth]{Figures/melt_dT_tav12.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 3.5cm 0 0},clip,width=\textwidth]{Figures/melt_dslope_tav12.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 3.5cm 0 0},clip,width=\textwidth]{Figures/gamma_T__e_res__dslope_tlim52.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 3.5cm 0 0},clip,width=\textwidth]{Figures/melt_dslope_tav12.png}
    \end{minipage}%    \caption{(a) Melt rate vs. thermal driving. (b) Melt rate vs. slope}
    \label{fig:melt_sensitivity}
\end{figure}
